---
title: "Descriptive Statistics in R"
author: "Nickie N."
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
if(!require(MASS)) install.packages("MASS")
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(Hmisc)) install.packages("Hmisc")
if(!require(vcd)) install.packages("vcd")
require(MASS)
library(MASS)
library(ggplot2)
require(ggplot2)
library(plyr)
library(lattice)
library(Hmisc)
library(vcd)
```

## Data exploration and basic visualization

### Import the data and get basic information pertaining to the columns

This data set contains data about 236 individuals. The columns relate to sex, exercise level, smoking habits, height, pulse, etc.

The first step was to remove any missing entries for sex as that is the basis for the visualizations below. 

```{r}
survey<-survey[!(is.na(survey$Sex)), ]
describe(survey)
```


### Boxpots
Boxplots were created for the following columns

* Pulse 
* Height  

#### Pulse

```{r}
ggplot(survey,aes(x="",survey$Pulse))+geom_boxplot(na.rm = TRUE)
```

#### Height

```{r}
ggplot(survey,aes(x="",survey$Height))+geom_boxplot(na.rm = TRUE)
```

### Review the five variable summary, mean, and standard deviation for the two variables above

#### Pulse

```{r}
summary(survey$Pulse)
sd(survey$Pulse, na.rm = TRUE)
mean(survey$Pulse, na.rm = TRUE)
```

#### Height

```{r}
summary(survey$Height)
sd(survey$Height, na.rm = TRUE)
mean(survey$Height, na.rm = TRUE)
```

## Describe Bivariate Relationships

###Determine if men and women have different smoking habits by creating a table.

```{r}
Smoke <- factor(survey$Smoke, levels = c("Never","Occas","Regul", "Heavy"))
table(survey$Sex, Smoke)

```

### Create a visualization of the relationship between sex and smoking behaviors.

```{r}
ggplot(data=survey,aes(Smoke))+geom_bar(na.rm = TRUE) + facet_grid(.~survey$Sex)

```
## Test the null hypothesis that pulse rate is similar across all exercise levels

### Create a box plot showing pulse rate grouped by exercise level 

```{r}
Exer2 <- factor(survey$Exer, levels = c("None", "Some", "Freq"))
ggplot(survey, aes(x = Exer2, y = Pulse)) +
  geom_boxplot()
```

### Conduct an analysis of variance to test the hypothesis

The *p*-value is less than 0.05, meaning that the hypothesis is rejected. There is a difference between pulse and exercise level. 

```{r}
ff <- aov(survey$Pulse~survey$Exer)
summary(ff)
```

## Line Graph using ggplot

### This plot shows unemployment over time. 

```{r}
ggplot(economics, aes(date, unemploy)) + geom_line()
```

### Update the plot to show the sitting president and their respective party.

```{r}
ggplot(economics) +
  geom_rect(
    aes(xmin = start, xmax = end, fill = party),
    ymin = -Inf, ymax = Inf, alpha = 0.2, data = presidential
    ) +
  geom_vline(
    aes(xintercept = as.numeric(start)),
    data = presidential,
    colour = "grey50", alpha = 0.5
    ) +
  geom_text(
    aes(x = start, y = 2500, label = name),
    data = presidential,
    size = 3, vjust = 0, hjust = 0, position = position_nudge (x = 50)
    ) +
  geom_line(aes(date, unemploy)) + scale_fill_manual(values = c("blue", "red")) 
```

## Session Info

```{r}
sessionInfo()
```
